Paper on the carbonate chemistry at Ny-Ålesund

### 2020-03-19 paper.Rmd
finalized temp and salinity

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
#library(robfilter)
library(seacarb)
library(gridExtra)
library(lubridate)
#library(lmtest)
#library(grid)
library(viridis)
#library(dygraphs)
require("knitr")
library("lmodel2")
#library(captioner)
#library(xts)
#library(seismicRoll)
library(scales)
library(plotly)
library(htmlwidgets)
library(tsibble)
if (!require("directlabels")) install.packages("directlabels")
library("directlabels")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

@@

# TEMP in situ SBE38
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_insitu), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="black", na.rm=TRUE, size=size_point) + 
  labs(title="Temperature SBE38",x="", y="In situ temperature (°C)") +
  #coord_fixed() +
  Mytheme(size_labs = 8)

@@

#Temperature vs salinity
We explore the use of a TS diagram to investigate changes in water sources. Now need to calculate slopes and r2, and see whether some months have a consistently different behaviour.

@@

d <- d_all %>%
  dplyr::filter(year < 2020) %>%
  dplyr::mutate(month=month(datetime, label = TRUE, abbr = TRUE)) # abbreviated months rather than digital months
  #ggplot(aes(x = sal_fb, y = temp_insitu), na.rm=TRUE) +
temp_sal <-  ggplot() +
  geom_point(data = d, aes(x = sal_fb, y = temp_insitu, fill=month), na.rm=TRUE,
             size=size_point, show.legend=FALSE) + 
  labs(title="Temperature vs salinity, blue lines show density", x="Salinity", y="Temp. (°C)") +
  facet_wrap(~ year, ncol = 3) +
  stat_smooth(data = d, aes(x = sal_fb, y = temp_insitu, color=month),
              method=lm, na.rm=TRUE, se=FALSE) +
  scale_color_viridis(name="Month", discrete=TRUE, option="plasma") +
  geom_contour(data = grid, aes(x = s, y = t, z = z), 
               stat = "contour", binwidth=1, show.legend=FALSE) +
  #stat_contour(aes(colour=..level..), breaks=seq(1025, 1028, by=1)) +
  coord_fixed() +
  Mytheme(size_labs = 10)
ggsave(temp_sal,filename = paste0(path_fig, "temp_sal.png"), width = 20, height = 20 ,units="cm")
temp_sal_py <- ggplotly(temp_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "temp_sal_py.html")
saveWidget(temp_sal_py, filename.html, selfcontained = T, libdir = "lib")
```

```{r temperature vs salinity regressions, fig.width=8, fig.height=8, echo=FALSE, message= FALSE}
# AT as a function of salinity for every year
j <- 0
reg <- data.frame()
for (i in 1:length(unique(d$year_month))) { #one loops through all years
  j <- j + 1
  ym <- unique(d$year_month)[i]
  d_reg <- filter(d, year_month==ym)
  if (length(!is.na(d_reg$temp_fb)) > 2) { #there is no temp_insitu in 2016-08, hence skip. Not needed if one uses temp_fb
  lms <- summary(lm(data=d_reg, sal_fb ~ temp_fb))
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <- pf(lms$fstatistic[1],lms$fstatistic[2],lms$fstatistic[3],lower.tail=FALSE)
  reg <- rbind(reg, as.numeric(c(lms$coefficients[2,1], lms$coefficients[2,2], 
                                 lms$coefficients[2,4], lms$coefficients[1,1], lms$coefficients[1,2],
                                 lms$coefficients[1,4], lms$fstatistic[1], lms$fstatistic[3], lms$r.squared, prob)))
  }
}
names(reg) <- c("Slope", "SE Slope", "P Slope","Intercept","SE int.","P int.", "F", "df",
                "R2","P value")
row.names(reg) <- as.character(unique(d$year_month))
#transform in latex table and save
tab_reg_temp_sal <- reg
tmp_reg_temp_sal <- reg %>% # to enable plotting below
  dplyr::mutate(date = as.Date(paste0(row.names(reg), "-01")),
                month = month(date, label=TRUE, abbr = TRUE))
#plot
fig_reg_temp_sal <- tmp_reg_temp_sal %>%
  ggplot() +
  geom_point(aes(x=date, y=Slope, color=month), size=5) +
  scale_color_viridis(name="Month", discrete = TRUE) +
  Mytheme(size_labs = 14)
fig_reg_temp_sal
```
kable(tab_fig_dyf_surf_date_AT)
print(dyf_surf_salinity_AT)

>>>>

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
#library(robfilter)
library(seacarb)
library(gridExtra)
library(lubridate)
#library(lmtest)
#library(grid)
library(viridis)
#library(dygraphs)
require("knitr")
library("lmodel2")
#library(captioner)
#library(xts)
#library(seismicRoll)
library(scales)
library(plotly)
library(htmlwidgets)
library(tsibble)
if (!require("directlabels")) install.packages("directlabels")
library("directlabels")
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

@@

# TEMP ferrybox
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_fb), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="black", na.rm=TRUE, size=size_point) + 
  labs(title="Temperature FerryBox",x="", y="In situ temperature (°C)") +
  #coord_fixed() +
  Mytheme(size_labs = 8)

@@

#Temperature vs salinity
We explore the use of a TS diagram to investigate changes in water sources. Monthly slopes of the T vs S plots are calculated. It seems that slopes are mostly not significant in winter and that they can be both positive and negative in summer.

@@

d <- d_all %>%
  dplyr::filter(year < 2020) %>%
  dplyr::mutate(month=month(datetime, label = TRUE, abbr = TRUE)) # abbreviated months rather than digital months
  #ggplot(aes(x = sal_fb, y = temp_insitu), na.rm=TRUE) +
temp_sal <-  ggplot() +
  geom_point(data = d, aes(x = sal_fb, y = temp_fb, fill=month), na.rm=TRUE,
             size=size_point, show.legend=FALSE) + 
  labs(title="Temperature vs salinity, both in FerryBox, blue lines show density", x="Salinity", y="Temp. (°C)") +
  facet_wrap(~ year, ncol = 3) +
  stat_smooth(data = d, aes(x = sal_fb, y = temp_fb, color=month),
              method=lm, na.rm=TRUE, se=FALSE) +
  scale_color_viridis(name="Month", discrete=TRUE, option="plasma") +
  geom_contour(data = grid, aes(x = s, y = t, z = z), 
               stat = "contour", binwidth=1, show.legend=FALSE, na.rm=TRUE) +
  scale_x_continuous(limits = c(30.4, 35.3), expand = c(0,0) ) +
  scale_y_continuous(limits = c(-2,8.5), expand = c(0,0) ) +
  coord_fixed() +
  annotate("text", label="Fresh and cold", x=31.5, y=-1.5) +
  annotate("text", label="Salty and warm", x=34.5, y=8.1) +
  Mytheme(size_labs = 10)
ggsave(temp_sal,filename = paste0(path_fig, "temp_sal.png"), width = 20, height = 20 ,units="cm")
temp_sal_py <- ggplotly(temp_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "temp_sal_py.html")
saveWidget(temp_sal_py, filename.html, selfcontained = T, libdir = "lib")
```
print(temp_sal)

```{r temperature vs salinity regressions, fig.width=6, fig.height=6, echo=FALSE, message= FALSE}
# AT as a function of salinity for every year
j <- 0
reg <- data.frame()
for (i in 1:length(unique(d$year_month))) { #one loops through all years
  j <- j + 1
  ym <- unique(d$year_month)[i]
  d_reg <- filter(d, year_month==ym)
  if (length(!is.na(d_reg$temp_fb)) > 2) { #there is no temp_insitu in 2016-08, hence skip. Not needed if one uses temp_fb
  lms <- summary(lm(data=d_reg, sal_fb ~ temp_fb))
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <- pf(lms$fstatistic[1],lms$fstatistic[2],lms$fstatistic[3],lower.tail=FALSE)
  reg <- rbind(reg, as.numeric(c(lms$coefficients[2,1], lms$coefficients[2,2], 
                                 lms$coefficients[2,4], lms$coefficients[1,1], lms$coefficients[1,2],
                                 lms$coefficients[1,4], lms$fstatistic[1], lms$fstatistic[3], lms$r.squared, prob)))
  }
}
names(reg) <- c("Slope", "SE_Slope", "P Slope","Intercept","SE int.","P int.", "F", "df",
                "R2","P_value")
row.names(reg) <- as.character(unique(d$year_month))
#transform in latex table and save
tab_reg_temp_sal <- reg
tmp_reg_temp_sal <- reg %>% # to enable plotting below
  dplyr::mutate(date = as.Date(paste0(row.names(reg), "-01")),
                month = month(date, label=TRUE, abbr = TRUE))
#plot
fig_reg_temp_sal <- tmp_reg_temp_sal %>%
  dplyr::filter(P_value < .05) %>% #eliminate when correlation not significant
  ggplot() +
  labs(title = "Slope ± 2 x SE", subtitle = "3 months with correlation coefficient not significant were eliminated") +
  geom_point(aes(x=date, y=Slope, color=month, size=P_value)) +
  scale_color_viridis(discrete = TRUE) +
  #scale_fill_continuous(name="P value slope") +
  geom_hline(yintercept = 0) + 
  scale_size_continuous(range = c(3,7)) +
  geom_errorbar(aes(x=date, ymin=Slope-2*SE_Slope, ymax=Slope+2*SE_Slope), width=.2) +
  Mytheme(size_labs = 14) +
  guides(colour = guide_legend("Month", override.aes = list(size=4)), size = guide_legend("P value slope"))
fig_reg_temp_sal
```
kable(tab_reg_temp_sal)
print(fig_reg_temp_sal)
